Secure Composition of Robust and Optimising Compilers

### 2023-07-08 popl/main.tex
RelWork MemSafety

This section compares existing work with the one presented in this paper.
To this end, work on robust preservation criteria (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is being looked at.
The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler--level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
This work has made use of that previous work.
While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
However, they do not demonstrate the scalability of the approach in the context of a compilation chain.

@@

There is still no consensus what the formal definition of memory safety as a program property should be. %cite
Yet, different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an attacker model that is not just the observing attacker.
For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of  linear typing~\cite{morrisett2005L3}.
The approach of this paper to memory safety and the countermeasures against memory safety insecurities are inspired by seminal work in the literature~\cite{nagarakatte2009soft,nagarakatte2010cets}.
The view of memory safety in this paper exhibits the classic split between temporal and spatial memory safety.
The former making sure that there are no double deallocations or use after frees.
The latter making sure that there are no out of bounds accesses.
When extending the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
The work presenting this monitor provides an approach for the robust {\em enforcement}, which implies robust preservation, of memory safety.
But, they do not compose the enforcement mechanism with other secure compilation passes.

>>>>

This section compares existing work with the one presented in this paper.
To this end, work on robust preservation (\Cref{subsec:relw:seccomprtp}) and on other secure compiler criteria (\Cref{subsec:relw:seccompcrit}) is looked at.
The case study of this paper (\Cref{sec:casestud:defs,sec:casestud:rtp}) implements measures for ensuring \gls{ms} as well as \gls{cct} and these are compared with previous work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler--level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2022universal,patrignani2019survey,kruse2022csc}.
This work has made use of that previous work.
While there exists a composition theorem~\cite{abate2019jour} already, that theorem is not concerned with composing robustly safe compilers, but rather program components and contexts.
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
However, they do not demonstrate the scalability of the approach by means of a case study.

@@

Different mechanisms for memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine--level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first--class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds--checks in front of pointer--dereferences and, for this to work, inserts meta--data information on pointer creation.
It also works in a more advanced setting with structured fields accesses and also introduces a table--lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects, i.e., there are no pointers to pointers or structures.
Of course, there are a lot of other approaches to memory--safety in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware--extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}, to name a few.
What differentiates this work from them is that this work uses known, compiler--based approaches to ensure memory--safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory--safety, but serves as a theoretical foundation for the secure compilation domain.

To extend the languages in this paper with a less restricted form of pointer arithmetic, the region coloring memory safety monitor presented in earlier work~\cite{michael2023mswasm} can be used.
The work presenting this monitor provides an approach for the robust preservation of memory safety compiling from C to WASM.
However, they do not discuss composition of secure compilers but rather investigate an instance of a secure compiler.
